---
title: "si618hw1_report_jlwohlf"
author: "John W"
date: "Tuesday, March 10, 2015"
output: html_document
---

Step 1: Load data

First the provided TSV data file is loaded into R using the read.table() function. Here are the first 15 rows
of the data frame:

```{r, echo=FALSE}
countrydata_withregion = read.delim("countrydata_withregion.tsv")
head(countrydata_withregion, n=15)
```

Step 2: Scatter plot of log transformed data

Natural logarithms of the area and the population of each country are computed and used to produce the
following scatter plot using the plot() function.

```{r, echo=FALSE}
plot(log(countrydata_withregion$area), log(countrydata_withregion$population))
```

Step 3: Data aggregation by region

The areas and populations of all countries in a region are summed up using the aggregate() function,
respectively. Then the following two pie charts are created using the pie() function.

```{r, echo=FALSE}
country.areas = aggregate(countrydata_withregion$area, by=list(countrydata_withregion$region), FUN=sum)

country.pop = aggregate(as.numeric(countrydata_withregion$population), by=list(countrydata_withregion$region), FUN=sum, na.rm=TRUE)

pie(country.areas$x, labels=country.areas$Group.1, main="Area of Regions")

pie(country.pop$x, labels=country.pop$Group.1, main="Population of Regions")
```
Step 4: Visualization of Population per sq km of Regions

A new data frame is created to contain the population per sq km of each region using the data.frame()
function. The data frame is then sorted by population per sq km in decreasing order with the help of the
order() function. Finally, the following bar plot is created using the barplot() function.

```{r, echo=FALSE}
pop_mile = data.frame(country=country.pop$Group.1, popm=(country.pop$x/country.areas$x))
mile = pop_mile[order(pop_mile$popm, decreasing = T),]
barplot(mile$popm, xlab="Population per sq km of Regions",names.arg=mile$country, width=2, cex.axis=1.5, cex.names=.5)
```